Identifying Heterogeneity in SAR Data with New Test Statistics

### 2024-12-14 Text/2024PresentationsChina.tex
improved and updated slides

\title{Identifying Heterogeneity in SAR Data with New Test Statistics}
\subtitle{No problem is too old, and you don't always need deep learning}
\author{Alejandro C.\ Frery}
\institute{Victoria University of Wellington\\
School of Mathematics and Statistics\\
New Zealand}
\date{November, 2024}

@@

\begin{frame}{The usual approach: I}
We are tempted to form a statistical test using the coefficient of variation as the test statistic.

Consider the random sample $\bm Z=(Z_1,Z_2,\dots,Z_n)$ that, under the null hypothesis has $\Gamma_{\text{SAR}}$ distribution.
Then, we need the distribution of
\begin{equation}
\widehat{\text{CV}}(\bm Z) = \frac{\frac1n \sum_{\ell=1}^n Z_\ell}{\sqrt{\frac1n \sum_{\ell=1}^n \big(Z_\ell-\overline{\bm Z}\big)^2}}.
\label{eq:CV}
\end{equation}

Although we know that $\overline{\bm X}$ and $\widehat{\text{CV}}(\overline X) $ are independent \citep{OnaCharacterizationoftheGammaDistributiontheIndependenceoftheSampleMeanandtheSampleCoefficientofVariation}, we don't know the distribution of~\eqref{eq:CV}.
\end{frame}

@@

Their estimator becomes
\begin{equation}
\widecheck{\text{CV}}(\bm Z) = \frac{\widehat{Q}_2(\bm Z)}{\frac1n \sum_{\ell=1}^n \big|Z_i-\widehat{Q}_2(\bm Z)\big|}.
\end{equation}

@@

Now, assume that we have a good estimator for the Shannon entropy, say $\widehat{H_{\Gamma_{\text{SAR}}}}$.
Under the $\Gamma_{\text{SAR}}$ model, we expect it to be close to~\eqref{E:E-gamma}.
Such estimator estimates
$$
\widehat{H_{\Gamma_{\text{SAR}}}} = \reallywidehat{L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) + \ln \mu}
$$

Now we build a test statistic:
\begin{align*}
	S(\bm Z; L) = & \hskip1ex\reallywidehat{L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) + \ln \mu} - \mbox{}\\
	& \hskip1ex L + \ln L - \ln\Gamma(L) - (1-L)\psi^{(0)}(L) - \ln \widehat{\mu}\\
	= & \hskip1ex \widehat{H_{\Gamma_{\text{SAR}}}} - L + \ln L - \ln\Gamma(L) - (1-L)\psi^{(0)}(L) - \ln \widehat{\mu}.
\end{align*}
This test statistic is close to zero under the null hypothesis (homogeneity, fully developed speckle, the $\Gamma_{\text{SAR}}$ model).

Now we must choose suitable estimators $\widehat{H_{\Gamma_{\text{SAR}}}}$ and $\widehat{\mu}$.
\end{frame}

@@

We are also interested in non-parametric alternatives.

@@

\begin{frame}{Finally\dots}
	We have three test statistics:
	\begin{align*}
		S_{\widehat{\text{CV}}}(\bm Z;L) &= \frac{\frac1n \sum_{\ell=1}^n Z_\ell}{\sqrt{\frac1n \sum_{\ell=1}^n \big(Z_\ell-\overline{\bm Z}\big)^2}} - \frac1{\sqrt{L}} \\
		S_{\widehat{\text{CV}_{\text{MnAD}}}} (\bm Z;L) &= \frac{\widehat{Q}_2(\bm Z)}{\frac1n \sum_{\ell=1}^n|X_i-\widehat{Q}_2(\bm Z)|}-\frac1{\sqrt{L}},\\
		S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L) & = \widetilde{H}_{\text{AO}}(\bm Z) - L + \ln L - \ln\Gamma(L) - (1-L)\psi^{(0)}(L) - \ln \widehat{\mu}
	\end{align*}

@@

We tested their size (error Type~I) and power (error Type~II) and finite-size distribution under the null hypothesis.
\alert{The winner is\dots}
\end{frame}

>>>>

\title{Identifying Heterogeneity in SAR Data with New Test Statistics}
\subtitle{No problem is too old, and you don't always need deep learning}
\author{Alejandro C.\ Frery}
\institute{Victoria University of Wellington\\
School of Mathematics and Statistics\\
New Zealand}
\date{December, 2024}

@@

\begin{frame}{The usual approach: I}
We are tempted to form a test statistic using the coefficient of variation.

Consider the random sample $\bm Z=(Z_1,Z_2,\dots,Z_n)$ that, under the null hypothesis has $\Gamma_{\text{SAR}}$ distribution.
Our test statistic is
\begin{equation}
S_{\widehat{\text{CV}}}(\bm Z;L) = \overbrace{\frac{\frac1n \sum_{\ell=1}^n Z_\ell}{\sqrt{\frac1n \sum_{\ell=1}^n \big(Z_\ell-\overline{\bm Z}\big)^2}}}^{\widehat{\text{CV}}(\bm Z)} - \underbrace{\frac{1}{\sqrt{L}}}_{\text{CV}\mid \Gamma_{\text{SAR}}} . \label{eq:SCV}
\end{equation}

Although we know that $\overline{\bm X}$ and $\widehat{\text{CV}}(\overline X) $ are independent \citep{OnaCharacterizationoftheGammaDistributiontheIndependenceoftheSampleMeanandtheSampleCoefficientofVariation}, we don't know the distribution of~\eqref{eq:SCV}.
\end{frame}

@@

Our robust test statistic becomes
\begin{equation}
S_{\widecheck{\text{CV}}}(\bm Z;L) = \frac{\widehat{Q}_2(\bm Z)}{\frac1n \sum_{\ell=1}^n \big|Z_i-\widehat{Q}_2(\bm Z)\big|} - \frac{1}{\sqrt{L}}.
\end{equation}

@@

Assume we have an estimator $\widehat{H}$ for the Shannon entropy of \alert{any} model.
Under the $\Gamma_{\text{SAR}}$ model, we expect it to be close to~\eqref{E:E-gamma}.
Such estimator estimates
$$
\widehat{H}\mid {\Gamma_{\text{SAR}}} (L,\mu) = \widehat{H_{\Gamma_{\text{SAR}}}}  = \reallywidehat{L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) + \ln \mu}
$$

Now we build a test statistic:
\begin{align*}
	S(\bm Z; L) = & \hskip1ex\reallywidehat{L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) + \ln \mu} - \mbox{}\\
	& \hskip1ex L + \ln L - \ln\Gamma(L) - (1-L)\psi^{(0)}(L) - \ln \widehat{\mu}\\
	= & \hskip1ex \widehat{H} - L + \ln L - \ln\Gamma(L) - (1-L)\psi^{(0)}(L) - \ln \widehat{\mu}.
\end{align*}
This test statistic is close to zero under the null hypothesis (homogeneity, fully developed speckle, the $\Gamma_{\text{SAR}}$ model).

Now we must choose suitable estimators $\widehat{H}$ and $\widehat{\mu}$.
\end{frame}

@@

@@

\begin{frame}{Finally\dots}
	We have three test statistics:
	\begin{align*}
		S_{\widehat{\text{CV}}}(\bm Z;L) &= \frac{\frac1n \sum_{\ell=1}^n Z_\ell}{\sqrt{\frac1n \sum_{\ell=1}^n \big(Z_\ell-\overline{\bm Z}\big)^2}} - \frac1{\sqrt{L}} \\
		S_{\widecheck{\text{CV}}}(\bm Z;L) &= \frac{\widehat{Q}_2(\bm Z)}{\frac1n \sum_{\ell=1}^n|X_i-\widehat{Q}_2(\bm Z)|}-\frac1{\sqrt{L}},\\
		S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L) & = \widetilde{H}_{\text{AO}}(\bm Z) - L + \ln L - \ln\Gamma(L) - (1-L)\psi^{(0)}(L) - \ln \widehat{\mu}
	\end{align*}

@@

We tested their size (error Type~I under the null hypothesis) 
and power (error Type~II under the $G_I^0$ model) and finite-size distributions.
\end{frame}
